---
title: "Superstore Data Analysis"
author: "Ayan Dey"
Date: "2023-10-24"
output: html_document
---

```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Superstore Data Analysis

Today I am going to work on Superstore data set. Analyse the data by \*\*R\*\* and get some useful insights about the data.

firstly set up the working directory

```{r}
# to see the current working directory
getwd()
```

```{r warning= TRUE}
# set up the working directory
setwd("C:\\Users\\growth\\Desktop\\R project")
```

```{r}
# confirmation of working directory
getwd()
```

```{r}
# importing all the  libraaries what I need the most 
 
library(tidyverse)
library(skimr)

```

```{r}
# importing the data -- I will always take the master dataset in variable = ** x **
x <- read.csv("superstore.csv")
x %>% head(7)
```

## Explore Understand and Clean the data

```{r}
# see the dimenstion of the dataset
dim(x)
```

Number of observation is : 9996

Number of variable is : 22

```{r}
# let see all the names of the 22 variables 
names(x)
```
```{r}
head(x$sub.category)
```


```{r}
# Rename sub.Category variable name into Sub_Category
x <-x %>% rename(sub_category=sub.category)

```

```{r}
# confirmation
names(x)
```

```{r}
# let rename all the variable name to smaller case
names(x) <- tolower(names(x))
```

```{r}
# confirmation
names(x)
```

```{r}
# let see the structure and datatype of variables of data
glimpse(x)
```

now work with each variable

```{r}
# change datatype of order_date variable datatype into - date datatype
x$order_date<- as.Date(x$order_date)
```

```{r}
# confirmation
class(x$order_date)
```

```{r}
# change datatype of ship_date variable datatype into - date datatype
x$ship_date<- as.Date(x$ship_date)
```

```{r}
# confirmation
class(x$ship_date)
```

```{r}
# change datatype of ship_mode variable datatype into factor datatype
x$ship_mode<- as.factor(x$ship_mode)
```

```{r}
# confirmation
class(x$ship_mode)
```

```{r}
# change datatype of sales variable datatype into - Numeric datatype
x$sales<- as.numeric(x$sales)
```

NULL value introduced in sales vector

```{r}
# let ignore the NULL value for now and do the pending operation
x$sales<- as.numeric(x$sales,na.rm=TRUE)
```

```{r}
# confirmation
class(x$sales)
```

```{r}
# let see again the structure of the data
glimpse(x)
```

-   there are a x variable which make me dought so let work on it

    ```{r}
    # let see the total length of x variable
    length(x$x)
    ```

```{r}
# there have total 9996 obs. 
# let see how many NULL value are there
sum(is.na(x$x))
```

```{r}
# let see how much data hoding that variable
sum( ! is.na(x$x))
```

total 9996 values and 9990 NULL

total 6 values are there

```{r}
# let locate those values
which( ! is.na(x$x))
```

```{r}
# let see all those obs. and those values understand importancy
x[ which (!is.na(x$x)), ] 
```

**conclusion** : I can't just drop the x variable becaue it hold number 6 values

but those values are unidentified and very confusion , what the values belongs to

after confirmation I decided to drop this variable.

```{r}
# now let locate the variable
which(names(x)== "x")
```

```{r}
# let remove that  by creating another variable ( Y)
y<-x[ , -22]
```

```{r}
# confirmation
names(y)
```

now working variable will be : [**y**]{.underline}

```{r}
print(" The new dimension is: " )
dim(y)
```

```{r}
# final confirmation 
glimpse(y)
```

all variable name , data type are set

now its time to deep dive inside the data

## NULL values

```{r}
# let check NULL values of each variable
# y <- working data
for ( i in 1:ncol(y)){
    print(paste(" null value of ",names(x[i]),"Obs. ", " index: ",i ))
     print(sum(is.na(y[i])) )
        }
```

let see each variable which have NULL values

```{r}
# let see all those data with max null
y[ , c(1:5)] %>% head(7)
```

```{r}
# let see with out order_date and ship_date
head(y[ , -c(3:4)])
```

It understood that this two column don't have much value right now to analysis , so I decided to drop this two columns .

-But also there have null in quantity(1), sales(6), discount(6) and profit ( 1) ,but I should not remove all those rows and just ignore when analysis.

```{r}
# let store the data without order_date and ship_date
y<- y[ , -c(3:4)]
```

```{r}
head(y)
```

#### Duplicates

```{r}
# just over view the boolean matrix 
# duplicated(x)
```

```{r}
# retrive all duplicate obs.
z<-y[duplicated(y), ]
```

```{r}
# let see the duplicate values
z
write.csv(z,"see_in_LibreOffice.csv")
```

```{r}
# now remove all the duplicate obs. and make data clean
y<-y[!duplicated(y), ]
head(y)
```

#### Now the data is cleaned and ready to save and Analysis**

```{r}
# now save the cleaned file 
# write.csv(y,"superstore_cleaned.csv")
```

## Analyse

```{r}
glimpse(y)
```

now I will work with each categorical vectors

##### Let Analyse the segment vector

```{r}
# to see the repeating factor values are in the ship_mode 
sort(table(y$ship_mode),decreasing = TRUE)
```

```{r}
# let see by graph
barplot(sort(table(y$ship_mode)))
```

##### Analyse the ship_mode vector with respect to quantity values

```{r}

#  for(i in 1 : nrow(y) ){
#   ifelse(y$ship_mode == 'Second Class', sum1[ , ]<-sum(y$quantity[i, ]),1*1 ) 
#     sum1= count(y$quantity[i, ])}
# }
# print(sum1) 
```

pull the ship_mode factore w.r.t frequency of quantity

```{r}
# pull with frequency with First class
ship_mode_result<- y %>%
  select(ship_mode,quantity) %>% 
  filter(ship_mode == "First Class")
m<-sum(ship_mode_result$quantity)
m
```

```{r}
# pull with frequency with Secound class

ship_mode_result<- y %>%
  select(ship_mode,quantity) %>% 
  filter(ship_mode == "Second Class")
n<-sum(ship_mode_result$quantity)
n
```

```{r}
# pull with frequency with Standard class

ship_mode_result<- y %>%
  select(ship_mode,quantity) %>% 
  filter(ship_mode == "Standard Class")
o<-sum(ship_mode_result$quantity,na.rm=TRUE)
o
```

```{r}
# pull with frequency with Same Day

ship_mode_result<- y %>%
  select(ship_mode,quantity) %>% 
  filter(ship_mode == "Same Day")
p<-sum(ship_mode_result$quantity,na.rm=TRUE)
p
```

```{r}
# let create a data frame of ship_mode for deep analyse
ship_class <-c ("First Class","Second Class","Standard Class","Same Day")
quantity <- c(m,n,o,p)
ship_mode_result <- data.frame( ship_class , quantity )
```

```{r}
# showing values
ship_mode_result
```

```{r}
# let create a pie chart to understand 
p<- c(ship_mode_result$quantity)
labels <- c(ship_mode_result$ship_class)
pie(p,labels,main="ship mode of each class", col= rainbow(length(x)))
```

Let Analyse category vector

```{r}
# let see the frequency
table(y$category)
```

```{r}
ggplot(data=y)+
  geom_bar(mapping=aes(x=category),fill="blue",color="red")+
  labs(title = "Frequency of category",
       subtitle="this bar graph shows which type of category are ordered")
```

### Now working with Numeric Data

Let work with sales vector

```{r}
class(y$sales)
```

ok : ( data Type )

```{r}
# let see over view of the distribution of sales vector
summary(y$sales)
```

```{r}
boxplot(y$sales)
```

as I see max data are contain within 17.248 to 209.858 , also have many outliers ( shown in box plot) and some NULL (6) values. dense

```{r}
# let take the maximun dense data
# And see the mean sales without outliers and NULL valuee 
mean_sales <- y %>% select(sales) %>% drop_na(sales) %>% 
  filter(sales >= 17.248 & sales <= 209.858)
mean(mean_sales$sales)
```

now compare it with median, which can give very closely result.

-   mean = 229.63

-   mean with out NULL & outliers = [**72.13**]{.underline}

-   median = [**54.356**]{.underline}

    **Time series**

```{r}
glimpse(x)
```

Let decide which variable is good option for Time series

```{r}
# let see at order_date variable
print(paste("total lenth : ",length(x$order_date)))
print(paste("total NULL values : ",sum(is.na(x$order_date))))
```

```{r}
# let see at ship_date variable
print(paste("total lenth : ",length(x$ship_date)))
print(paste("total NULL values : ",sum(is.na(x$ship_date))))
```

ship_date (NULL 6097 ) \> order_date(NULL 5952)

though there is no cancel order . it is good idea to work with order_date.


#### Now do some Question Answer solve

```{r}
## which are the most selling product ?
x1<-y %>% mutate('total_sales' =(sales*quantity)-(discount*quantity))
x1 %>% arrange(-total_sales) %>% select(product_name) %>% head(10)
```
